---
title: "Homework 2"
author: Michelle Lui
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr. Trashwheel dataset. 

```{r}
trashwheel_df = 
    read_xlsx(
      "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
      sheet = "Mr. Trash Wheel",
      range = cell_cols("A:N")) %>%
    janitor::clean_names() %>%
    drop_na(dumpster) %>%
  mutate(
      sports_balls = round(sports_balls),
      sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data! For 2018 and 2017.

```{r}
precip_2018 = 
    read_excel(
      "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
      sheet = "2018 Precipitation",
      skip = 1
    ) %>%
    janitor::clean_names()%>%
    drop_na(month) %>%
    mutate(year = 2018) %>%
    relocate(year)

precip_2017 = 
    read_excel(
      "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
      sheet = "2017 Precipitation",
      skip = 1
    ) %>%
    janitor::clean_names()%>%
    drop_na(month) %>%
    mutate(year = 2017) %>%
    relocate(year)
```

Now combine annual precipitation.

```{r}
month_df = 
  tibble(
      month = 1:12,
      month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017) 

left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters, the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. In this dataset:

* The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`.
* The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.

## Problem 2

Read the NYC Transit dataset.

```{r}
transit_df = 
    read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
    janitor::clean_names() %>%
    select(line:entry, vending, ada) %>%
    mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
  

```

This dataset contains the variables: line, station name, station latitude/longitude, routes served, entry, vending, entrance type, and ADA compliance. So far, for data cleaning steps I have cleaned the variable names, selected the desired variables from the dataset, and converted the entry variable from character to a logical variable. In this dataset there are `r nrow(transit_df)` rows and `r ncol(transit_df)` columns. The data is not yet tidy however because the routes variable is spread across 11 columns, and instead could be collapsed into one column, and the variable names for the routes contain values (1-11). There are `r distinct(transit_df, line, station_name) %>% count()` distinct stations in this dataset. There are `r filter(transit_df, ada ==TRUE) %>% distinct(line, station_name) %>% count()` stations that are ADA compliant. The proportion of station entrances / exits without vending that allow entrance is `r (filter(transit_df, vending == "NO") %>% filter(entry == TRUE) %>% count()) / (filter(transit_df, vending == "NO") %>% count())`. 

Reformat and tidy data.

```{r}
transit_tidy_data = 
    transit_df %>%
    mutate_at(
        vars("route1":"route11"),
        as.character
         ) %>%
    pivot_longer(
      route1:route11,
      names_to = "route_number",
      values_to = "route_name"
    ) %>%
    drop_na(route_name)
```

`r filter(transit_tidy_data, route_name == "A") %>% distinct(line, station_name) %>% count()` distinct stations serve the A train. Of the stations that serve the A train, `r filter(transit_tidy_data, route_name == "A") %>% filter(ada == TRUE) %>% distinct(line, station_name) %>% count()` are ADA compliant.
